\section{What we would like to do}

\begin{frame}{Implement more advanced controllers}

    Nonlinear controllers are particularly interesting because of the highly nonlinear nature of the system.
    However, they look like they are going to be a challenge to implement.

    \vspace{9pt}

    \begin{columns}[T, onlytextwidth]

        \begin{column}{0.5\textwidth}

            \textbf{Linear Controllers}:

            \begin{itemize}
                \item Model Predictive Control
                \item Cascaded Control
                \item Pole Placement
                \item Linear Quadratic Integrator
            \end{itemize}

        \end{column}

        \begin{column}{0.5\textwidth}

            \textbf{Nonlinear Controllers}:

            \begin{itemize}
                \item Backstepping
                \item Feedback Linearization
            \end{itemize}

        \end{column}

    \end{columns}

    \vspace{9pt}

    When possible, we would also like to implement logics such as Kalman Filters or gain scheduling.

\end{frame}



\begin{frame}{Study the effects of approximations}

    Once the characterization of the inductances is complete, we would like to study the effects of the approximations we have made in the model.

    \vspace{9pt}

    For example by comparing:

    \begin{itemize}
        \item Linearized vs nonlinear model with respect to the real world system;
        \item Effects of simplified ($L(z)$) vs complete ($L(z, I)$) inductance modelling.
    \end{itemize}

\end{frame}



\begin{frame}{Controllers comparison}

    We would like to compare the performance of the controllers we have implemented.

    \vspace{9pt}

    To do so, we will probably create a `Race of Controllers\footnotemark[1]' where we will compare various performance indices.

    \footnotetext[1]{More on this in the final presentation.}

\end{frame}