\section{Identification}
\label{sec:identification}

To effectively control the system, it is crucial to identify its physical and dynamic parameters.
This process involves a series of carefully designed experiments and measurements to extract reliable data.

The parameter identification process is divided into the following steps:

\begin{enumerate}
    \item \textbf{Direct measurement}: using conventional instruments, measure physical parameters that do not require specific test setups, such as geometric dimensions or mass or static resistance of the coils.
    \item \textbf{Sensor characterization}: map the relationship between the ball's position and the output voltage of the infrared sensor. Additionally, statistically analyze the output of the sensors used internally to estimate the system state.
    \item \textbf{Control-to-voltage dependency}: examine the relationship between the control signal and the resulting applied voltage to the coils ($V = V(U)$).
    \item \textbf{Inductance characterization}: determine parameters for inductances based on the model described in Equations \ref{eq:model_for_inductance}. This involves analyzing the system's electrical response under various conditions.
    \item \textbf{Force analysis}: measure the electromagnetic force applied to the ball to validate both the identified parameters and the overall model's accuracy;
    \item \textbf{Active levitation}: a preliminary active controlled levitation test to further refine the identified parameters.
\end{enumerate}

Except for direct measurements, all tests are performed using the data acquisition capabilities of the \texttt{Inteco} control unit itself.

To simplify the identification process, we will assume the lower and upper coils have identical parameters unless explicitly stated otherwise.
This assumption allows us to streamline both the methodology and the notation by avoiding subscripts that distinguish the two coils.

In all subsequent tests, parameters are identified using measurements from the upper coil.

\input{src/04.1 - direct_measurement.tex}
\input{src/04.2 - sensors_characterization.tex}
\input{src/04.3 - control_to_voltage.tex}
\input{src/04.4 - inductances_characterization.tex}
\input{src/04.5 - force_analysis.tex}
\input{src/04.6 - active_levitation.tex}
